\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.75in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 1410}\chead{Worksheet \# 11}\rhead{Tuesday, 5\textsuperscript{th} April, 2016}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newenvironment{amatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|c@{}}
}{%
  \end{array}\right]
}
\newenvironment{aamatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|*{#1}{c}@{}}
}{%
  \end{array}\right]
}

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}       
\DeclareMathOperator{\proj}{proj}   
\newcommand{\bam}{\begin{amatrix}}
\newcommand{\eam}{\end{amatrix}}
         
                  
\begin{document}

%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}

 \begin{enumerate}
  \item For the following matrices, find (i) the characteristic polynomial, (ii) the eigenvalues of the matrix, and (iii) the corresponding eigenvectors.
\begin{multicols}{3}

\begin{enumerate}
  \item $A = \bbm 3& 1&2\\0&2&-4\\0& -1&-1\ebm$
  \item $B = \bbm 3 & 1 & 2\\0& 2 & 1\\0& 0& 2\ebm$
  \item $C = \bbm 0&1&1\\1&0&1\\1&1&0\ebm$
 \end{enumerate}
\end{multicols}
\item For each of the matrices in problem 1, find an invertible matrix $P$ such that $P^{-1}AP$ is diagonal, or explain why no such $P$ exists. (You don't have to compute $P^{-1}$, unless you want to make sure you did everything correctly.)
%\item Suppose that the $n\times n$ matrix $A$ is diagonalizable, and we know that $\lambda^3 = 5\lambda$ for every eigenvalue $\lambda$ of $A$. Prove that $A^3=5A$.

%\textit{Hint:} since $A$ is diagonalizable, there exists a diagonal matrix $D$ and an invertible matrix $P$ such that $A=PDP^{-1}$.

\item An $n\times n$ matrix $A$ is called \textbf{orthogonal} if $A^T=A^{-1}$ (that is, $A^TA=I$).
\begin{enumerate}
 \item Show that the matrix $A=\bbm -1/3&2/3&2/3\\2/3&-1/3&2/3\\2/3&2/3&-1/3\ebm$ is orthogonal. 
 \item Prove that a matrix $A$ is orthogonal if and only if the columns of $A$ form an orthonormal set of vectors. (That is, the columns $C_1,\ldots, C_n$ of $A$ satisfy $\len{C_i}=1$ for each $i=1,\ldots, n$, and $C_i\dotp C_j = 0$ for each $i\neq j$.) 
\end{enumerate}

\item (\textbf{Bonus fun:}) An $n\times n$ matrix $A$ is called \textit{symmetric} if $A^T=A$. An important theorem in linear algebra, called the \textit{Spectral Theorem}, guarantees that every symmetric matrix $A$ can be ``orthongonally diagonalized'', meaning that there exists an \textbf{orthogonal} matrix $P$ such that $P^TAP = D$ is diagonal. (Note from the previous problem that $P^T=P^{-1}$.)
\begin{enumerate}
 \item Show that the matrix $A = \bbm 3&0&0\\0&2&2\\0&2&5\ebm$ is symmetric.
 \item It is possible to prove that eigenvectors corresponding to \textbf{distinct} eigenvalues of a symmetric matrix are orthogonal. Use this fact to find an orthogonal matrix $P$ such that $P^TAP$ is diagonal.
 \item Show that the matrix $C$ from problem 1(c) is also symmetric. You should have found that $C$ had only two eigenvalues: $\lambda =2$, and $\lambda = -1$ (which is repeated). Letting $X$ denote the eigenevector corresponding to $\lambda=2$, and letting $Y_1, Y_2$ denote the eigenevectors corresponding to $\lambda = -1$, show that $X$ is orthogonal to both $Y_1$ and $Y_2$.
 \item Chances are the two eigenvectors $Y_1$ and $Y_2$ corresponding to $\lambda = -1$ were not themselves orthogonal. Can you replace $Y_2$ by an eigenvector for $\lambda = -1$ that \textbf{is} orthogonal to $Y_1$? (\textit{Hint:} projection.)
\end{enumerate}

 \end{enumerate}

\newpage

{\large \bf Name:}

\vspace{24pt}

{\large \bf Tutorial time:}

\vspace{24pt}

Please submit \textbf{one} \textit{completed} solution from the worksheet for feedback. 

\textbf{Note:} Your solution needs to contain enough detail for it to be clear what problem you're trying to solve!

\end{document}