\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 1410}\chead{Worksheet \# 4}\rhead{Tuesday, 2\textsuperscript{nd} February, 2016}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}       
\DeclareMathOperator{\proj}{proj}            
                  
\begin{document}
{\bf \large Name:} \hspace{2.5in} {\bf Tutorial time:}

\bigskip


%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}
Please complete all problems below.
 \begin{enumerate}
 \item Solve the system of linear equations below by (a) forming the corresponding augmented matrix, and (b) using Gaussian elimination (row operations) to reduce it to row-echelon form.
\[
 \begin{array}{ccccccc}
  x&-&2y&+&z&=&4\\
 -x&+&y&-&2z&=&-2\\
 2x&-&4y&+&3z&=&9
 \end{array}
\]

\newpage

 \item Consider the following system of two equations in three variables:
\[
 \begin{array}{ccccccc}
  x&+&2y&-&3z=6\\
 2x&+&5y&-&4z=-3
 \end{array}
\]
\begin{enumerate}
 \item What geometric object in $\R^3$ (3-dimensional space) is defined by each of the individual equations? (A point? A line? A plane? Something else?)

\vspace{0.5in}

 \item Find a one-parameter family of solutions to the system of equations above. (It should only take one row operation.)

\vspace{2.5in}

 \item Write your solution to part (b) in vector form ( as $\bbm x\\y\\z\ebm = \ldots$). What geometric object in $\R^3$ does your solution represent? Does this seem reasonable?\\ (Note that any point $(x,y,z)$ that satisfies {\bf both} of the original equations belongs to the {\bf intersection} of the objects defined by those equations.

\vspace{1.5in}

\item Suppose it had turned out that the system of equations had no solutions. How would you explain this visually?
\end{enumerate}




 \end{enumerate}
\newpage

{\bf Exam preparation:} Do not hand in this page. Take it home with you, and make sure you can confidently answer every question below. (You will have to be able to do some -- perhaps all -- of these things on the test, so anything left unanswered represents marks left unearned on your test.)

\medskip

{\bf How do I ...}
\begin{enumerate}
 \item Compute the distance between two points? (In $\R^2$? In $\R^3$? In $\R^n$?)
 \item Determine the vector $\overrightarrow{PQ}$ whose tail is at the point $P$ and whose head is at the point $Q$?
 \item Determine the point $Q$, if given $P$ and $\overrightarrow{PQ}$, or the point $P$, if given $Q$ and $\overrightarrow{PQ}$?
 \item Calculate the length of a vector?
 \item Add two vectors together, or multiply a vector by a scalar?
 \item Calculate the dot product of two vectors? And use it to find the angle between two vectors?
 \item Define a line in $\R^3$? (What are the parametric equations? What is the vector equation?)
 \item Find the equation(s) of a line, given a point on the line and a vector parallel to the line (direction vector)? What if I'm given two points on the line? Or a point on the line and the equation(s) of a parallel line?
 \item Find either a point on a line, or a vector in the direction of the line, if given the equations of the line?
 \item Determine the equation of a plane, if given a point on the plane and the normal vector?
 \item Determine a point on a plane, or the normal vector to the plane, if given the equation of the plane?
 \item Compute a projection? And describe a projection using a diagram? (If I'm projecting a vector $\vec{a}$ onto a vector $\vec{b}$, and drawing each vector as an arrow, where do I locate the tails of $\vec{a}$, $\vec{b}$, and $\proj_{\vec{b}}\vec{a}$? Where do I locate the heads?)
 \item Take a vector $\vec{a}$, and, with resepct to a second vector $\vec{b}$, write it as $\vec{a} = \vec{a}_\parallel + \vec{a}_\bot$ where $\vec{a}_\parallel$ is parallel to $\vec{b}$, and $\vec{a}_\bot$ is orthogonal to $\vec{b}$?
 \item Find the shortest distance from a point to a line? Or a plane?
 \item Find the intersection of a line and a plane?
 \item Find the intersection of two lines?
 \item Find the intersection of two planes?
 \item Tell if a matrix is in (reduced) row-echelon form?
 \item Determine the solution of a system from the row-echelon form of its augmented matrix?
 \item Tell if system of equations has one solution, no solution, or infinitely many solutions?
\end{enumerate}
(Note that the last two items require you to solve a system of equations.)
\end{document}