\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 1410}\chead{Worksheet \# 5}\rhead{Tuesday, 9\textsuperscript{th} February, 2016}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newenvironment{amatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|c@{}}
}{%
  \end{array}\right]
}
\newenvironment{aamatrix}[1]{%
  \left[\begin{array}{@{}*{#1}{c}|cc}
}{%
  \end{array}\right]
}

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}       
\DeclareMathOperator{\proj}{proj}   
\newcommand{\bam}{\begin{amatrix}}
\newcommand{\eam}{\end{amatrix}}
         
                  
\begin{document}
{\bf \large Name:} \hspace{2.5in} {\bf Tutorial time:}

\bigskip

{\bf Problem you want feedback on:}

\bigskip

%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}
Please complete all problems below.
 \begin{enumerate}
 \item When you put the symbol ``='' between two objects on the page, what are you saying about the relationship between those objects?

\vspace{0.5in}

 \item Each of the augmented matrices below is in reduced row-echelon form. For each matrix, indicate the following:

 \begin{enumerate}
  \item The {\em rank} of the augmented matrix.
  \item The number of variables in the corresponding system of equations.
  \item The number of parameters needed to write down the general solution.
  \item The general solution to the system, if any.
\begin{multicols}{2}
\begin{enumerate}
 \item $\di \bam{3} 1&0&-2&4\\0&1&3&-5\\0&0&0&0\eam$

\vspace{2in}

 \item $\di \bam{4} 1&-3&0&4&2\\0&0&1&-3&7\\0&0&0&0&0\eam$
 \item $\di \bam{3} 1&2&0&3\\0&0&1&4\\0&0&0&1\eam$

\vspace{2in}

 \item $\di \bam{4} 0&1&0&-1&4\\0&0&1&0&2\\0&0&0&0&0\eam$
\end{enumerate}
\end{multicols}
 \end{enumerate}

\newpage

\item Determine the value(s) of $a$ such that the system of equations given by the augmented matrix below has no solution, one solution, or infinitely many solutions, if possible.
\[
\begin{amatrix}{3}
a & 1 & 2 & 1 \\

2 & 1 & 7 & 3 \\

1 & 1 & 2 & 1 \\
\end{amatrix}
\]

\vspace{3.75in}


\item Find the basic solutions to the homogeneous system of equations
\[
 \begin{array}{ccccccccc}
  x_1&-&2x_2&+&x_3&-&3x_4&=&0\\
 2x_1&+&x_2&-&4x_3&+&x_4&=&0\\
 3x_1&-&x_2&-&3x_3&-&2x_4&=&0
 \end{array}
\]



 \end{enumerate}

\end{document}