\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{Page \thepage \ of \pageref{LastPage}}\rfoot{{\bf Total Points: 40}}
\chead{MATH 4310}\lhead{Term Test II}\rhead{Monday, 17\textsuperscript{th} November, 2014}

\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\abs}[1]{\lvert #1\rvert}

\begin{document}

\author{Instructor: Sean Fitzpatrick}
\thispagestyle{plain}
\begin{center}
\emph{University of Lethbridge}\\
Department of Mathematics and Computer Science\\
17\textsuperscript{th} November, 2014, 5:00-5:50 pm\\
{\bf Math 4310 - Term Test II}\\
\end{center}
\skipline \skipline \skipline \noindent \skipline
Last Name:\underline{\hspace{350pt}}\\
\skipline
First Name:\underline{\hspace{348pt}}\\
\skipline
Student Number:\underline{\hspace{322pt}}\\


\vspace{0.5in}


\begin{quote}
 {\bf Record your answers below each question in the space provided.    Left-hand pages may be used as scrap paper for rough work.  If you want any work on the left-hand pages to be graded, please indicate so on the right-hand page.
 
 \bigskip
 
Partial credit will be awarded for partially correct work, so be sure to show your work, and include all necessary justifications needed to support your arguments. 
}
\end{quote}


\vspace{0.5in}

For grader's use only:

\begin{table}[hbt]
\begin{center}
\begin{tabular}{|l|r|} \hline
Page&Grade\\
\hline \hline
\cline{1-2} 2 & \enspace\enspace\enspace\enspace\enspace\enspace/12\\
\cline{1-2} 3 & \enspace\enspace\enspace\enspace\enspace\enspace/8\\
\cline{1-2} 4 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} 5 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} Total & \enspace\enspace\enspace\enspace\enspace\enspace/40\\
\hline
\end{tabular}

\skipline

\skipline

\skipline


\end{center}
\end{table}
\newpage


\begin{enumerate}
\item Let $X$ be a topological space. List as many conditions as possible equivalent to the statement ``$X$ is not connected.'' \points{6} You don't need to justify your responses.

(Note: It is possible to earn bonus points on this problem. I will give 2 points per correct response, up to 6 points, plus 1 point for any additional correct responses.)

\vspace{3.5in}

\item Let $X$ be a subspace of $\R^n$ (with the Euclidean topology). List as many conditions as possible equivalent to the statement ``$X$ is compact.'' \points{6} The same scoring rules apply as in Problem \#1.

\newpage

\item Let $X$ and $Y$ be topological spaces, where $X$ is compact and $Y$ is Hausdorff, and let $f:X\to Y$ be continuous.
\begin{enumerate}
 \item Prove that if $f:X\to Y$ is a surjection, then $f$ is a quotient map. \points{5}

\vspace{3.5in}

 \item Prove that if $f:X\to Y$ is a bijection, then $f$ is a homeomorphism. \points{3}



\end{enumerate}
\newpage
 \item Prove that if $f:X\to Y$ is a continuous surjection and $X$ is path-connected, then $Y$ is path-connected. \points{5}

\vspace{3.5in}

 \item Suppose that $X=A\cup B$, with $A,B$ nonempty, disjoint, open subsets of $X$. (That is, $\{A,B\}$ is a separation of $X$.) Prove that a map $f:X\to Y$ is continuous if and only if the restrictions $f|_A:A\to Y$ and $f|_B:B\to Y$ are continuous. \points{5}

\newpage

 \item Suppose $\{F_n | n\in\N\}$ is a family of nonempty closed subsets of a topological space $X$, such that $F_{n+1}\subseteq F_n$ for each $n\in \N$. Prove that if $X$ is compact, then $\bigcap_{n\in\N}F_n$ is nonempty. \points{6}

Hint: To get a contradiction, suppose $\bigcap_{n\in\N}F_n = \emptyset$, and take complements.

 \vspace{4.5in}

 \item Consider the cylinder $X=S^1\times [-1,1]$, and let $\sim$ be the equivalence relation on $X$ defined by the partition  $\mathcal{P}$ consisting of the sets $S^1\times\{-1\}$, $S^1\times\{1\}$, and the single point sets $\{(x,t)\}$, where $x\in S^1$ and $t\in (-1,1)$. Describe the quotient space $X/\sim$. \points{4} (A formal proof is not required.)


\end{enumerate}



\end{document}