\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 2560}\chead{Worksheet \# 4}\rhead{Thursday 11\textsuperscript{th} February, 2016}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}                   
                  
\begin{document}


%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}
The problems on this worksheet are for in-class practice during tutorial. You are free to collaborate and to ask for help. They don't count for course credit, but it's a good idea to make sure you know how to do everything before you leave tutorial -- similar problems may show up on a test or assignment.

\bigskip

Note: your solutions to most of these problems are not going to fit on this page. (I hope you brought paper with you.)

\begin{enumerate}
 \item For each of the following sets of curves, sketch the curves to determine the area they enclose, and then find the area.
\begin{enumerate}
 \item $y=\sqrt{x+2}$, $y=\dfrac{1}{x+1}$, $x=0$ and $x=2$.
 \item $y=2x^2+5x-3$ and $y=x^2+4x-1$.
 \item $y=x^2+1$, $y=\frac{1}{4}(x-3)^2+1$, and $y=1$.
 \item $y=\cos x$ and $y=\sin 2x$, between $x=0$ and $x=\pi/2$.
 \item $y=x$ and $y=x^3$.
 \item $y=x$, $y=5x$, and $y=6-x^2$, in the first quadrant. 
\end{enumerate}
\item Use calculus to find the following volumes:
\begin{enumerate}
 \item Of the solid $S$ whose base is the region of the $xy$-plane bounded by $y=x^2$ and $y=2-x^2$, and whose cross-sections parallel to the $y$-axis are squares.
 \item Of a right circular cone whose height is 12 and base radius is 4.
 \item The region bounded by $y=4-x^2$ and $y=0$, when rotated about
\begin{enumerate}
 \item the $x$-axis.
 \item the line $y=-1$.
 \item the line $x=2$.
\end{enumerate}
 \item The region bounded by the curves $y^2=x$ and $x=2y$, when revolved around the $y$-axis.
\end{enumerate}

\end{enumerate}


\end{document}