\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 2560}\chead{Worksheet \# 5}\rhead{Thursday 25\textsuperscript{th} February, 2016}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}                   
                  
\begin{document}


%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}
The problems on this worksheet are for in-class practice during tutorial. You are free to collaborate and to ask for help. They don't count for course credit, but it's a good idea to make sure you know how to do everything before you leave tutorial -- similar problems may show up on a test or assignment.

\bigskip

Some integration formulas you may need:

\begin{itemize}
 \item Volume of a solid of revolution, washer method, rotation about a horizontal axis: 
\[
 V = \pi\int_a^b (r_{out}(x)^2-r_{in}(x)^2)\,dx,
\]
where $r_{out}$ gives the outer radius (distance from the far side of the region being rotated to the axis) and $r_{in}$ gives the inner radius. If the axis of rotation is vertical, reverse the roles of $x$ and $y$.



 \item Volume of a solid of revolution, shell method, rotation about a vertical axis:
\[
 V = 2\pi\int_a^b r(x)h(x)\,dx,
\]
where $r(x)$ is the radius of the shell (distance to axis of rotation) and $h(x)$ is the height of the shell. If the region lies between $y=g(x)$ (above) and $y=f(x)$ (below), and the axis of rotation is the $y$-axis we get the formula
\[
 V = 2\pi \int_a^b x(g(x)-f(x))\,dx
\]
as a special case. If the axis of rotation is horizontal, reverse the roles of $x$ and $y$.

\item Arc length of a curve $y=f(x)$, for $a\leq x\leq b$:
\[
 L = \int_a^b\sqrt{1+f'(x)^2}\,dx.
\]
\item Surface area generated by revolving $y=f(x)$, $a\leq x\leq b$ about the $x$-axis:
\[
 S = 2\pi\int_a^b f(x)\sqrt{1+f'(x)^2}\,dx.
\]
\item Surface area generated by revolving $y=f(x)$, $a\leq x\leq b$ about the $y$-axis:
\[
 S = 2\pi\int_a^b x\sqrt{1+f'(x)^2}\,dx.
\]

\end{itemize}

\subsection*{Exercises}
\begin{enumerate}
 \item Find the volume of the solid generated by revolving the region bounded by $y=x^2$, $x=1$, and $y=0$ about the $x$-axis.
 \item Find the volume of the solid generated by revolving the region bounded by $y=\sqrt{x}$, $y=1$, and $x=0$ about the $x$-axis.
 \item Repeat Problem 2, but revolving about the $y$-axis.
 \item Find the volume of the solid generated by revolving the region bounded by $x=y-y^2$ and $x=0$ about the $y$-axis.
 \item Use the shell method to find the volume of the solid generated by revolving the region bounded by $y=6x-2x^2$ and $y=0$, about the $y$-axis.
 \item Use the shell method to find the volume of the solid generated by revolving the region bounded by $y=x$ and $y=\sqrt{x}$ about the $x$-axis.
 \item Find the length of the curve $y=\dfrac{1}{12}x^3+\dfrac{1}{x}$, for $x\in [1,4]$.
 \item Find the area of the surface obtained by revolving $y=\sqrt{x}$, for $x\in [0,1]$, about the $x$-axis.
 \item Find the area of the surface obtained by revolving $y=x^2$, for $x\in [0,1]$, about the $y$-axis.
 \item Find the area of the surface obtained by revolving $x=1+2y^2$, $1\leq y\leq 2$, about the $x$-axis.
\end{enumerate}


\end{document}