\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 2560}\chead{Worksheet \# 9}\rhead{Thursday 24\textsuperscript{th} March, 2016}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}                   
                  
\begin{document}


%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}
The problems on this worksheet are for in-class practice during tutorial. You are free to collaborate and to ask for help. They don't count for course credit, but it's a good idea to make sure you know how to do everything before you leave tutorial -- similar problems may show up on a test or assignment.

\begin{enumerate}
 \item Solve the following separable differential equations:
\begin{enumerate}
 \item $xy' = y+2x^2y$, where $y(1)=1$.
 \item $\dfrac{dy}{dx} = \dfrac{x^2+1}{y^2+1}$, where $y(0)=1$. (Give an implicit solution.)
 \item $(4y+yx^2)\,dy - (2x+xy^2)\,dx = 0$
\end{enumerate}
 \item Solve the following linear differential equations. State an interval on which the general solution is defined.
\begin{enumerate}
 \item $\dfrac{dy}{dx}+y=e^{3x}$
 \item $(1+x^2)\,dy+(xy+x^3+x)\,dx = 0$
 \item $(1-x^3)\dfrac{dy}{dx}=3x^2y$
 \item $(x^2+x)\,dy+(xy+x^3+x)\,dx = 0$
 \item $\cos x\dfrac{dy}{dx}+y\sin x =1$
\end{enumerate}
\item Determine whether the sequence converges or diverges. If it converges, give the limit.
\begin{enumerate}
 \item $a_n = (-1)^n\dfrac{n}{n^2+1}$
 \item $a_n = (-1)^n\dfrac{2n+1}{3n+4}$
 \item $a_n = \dfrac{n-1}{n}-\dfrac{n}{n-1}$
 \item $a_n = \dfrac{4n}{\sqrt{9n^2+4}}$
 \item $a_n = 1-\dfrac{1}{n}$
 \item The sequence $\{a_n\}$ defined by $a_1=1$ and $a_{n+1} = 3-\dfrac{1}{a_n}$ for all $n\geq 1$. (You may assume that the sequence converges. If you want to actually \textit{show} that it coverges, feel free to ask me how.
 \item The sequence $\{a_n\}$ defined by $a_1 = \sqrt{2}$ and $a_{n+1} = \sqrt{2+a_n}$ for all $n\geq 1$. (Again, you may assume that the sequence converges.)
\end{enumerate}



\end{enumerate}





\end{document}