\documentclass[letterpaper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}

\newcommand{\abs}[1]{\lvert #1\rvert}

\title{Math 1410 Lyryx Lab 2\\University of Lethbridge, Spring 2015}
\author{Sean Fitzpatrick}
\begin{document}
 \maketitle


\begin{enumerate}
 \item Solve the following system of equations:
\[
 \begin{array}{ccccccccc}
  2x_1&+&2x_2&-&8x_3&+&6x_4&=&-12\\
  -2x_1& & &+&8x_3&-&6x_4&=&8\\
x_1& & &-&4x_3&+&3x_4&=&-4
 \end{array}
\]
 \item Solve the following homogeneous system of linear equations. If the system has no solution, demonstrate this by giving a row-echelon form of the augmented matrix of the system.
\[
 \begin{array}{ccccccc}
  5x_1&-&5x_2&-&5x_3&=&0\\
x_1&-&x_2&-&x_3&=&0\\
-2x_1&+&3x_2&+&5x_3&=&0
 \end{array}
\]
\item Solve the following homogeneous system of linear equations. If the system has no solution, demonstrate this by giving a row-echelon form of the augmented matrix of the system.
\[
  \begin{array}{ccccccc}
 2x_1&+&8x_2&+&8x_3&=&0\\
2x_1&+&8x_2&+&8x_3&=&0\\
-x_1&-&4x_2&-&4x_3&=&0
 \end{array}
\]

\item Solve the following homogeneous system of linear equations. If the system has no solution, demonstrate this by giving a row-echelon form of the augmented matrix of the system.
\[
 \begin{array}{ccccccccc}
2x_1&+&8x_2&+&6x_3&+&4x_4&=&0\\
x_1&+&4x_2&+&6x_3&+&5x_4&=&0\\
-2x_1&-&8x_2&-&7x_3&-&2x_4&=&0\\
-2x_1&-&8x_2&-&4x_3&-&3x_4&=&0  
 \end{array}
\]
\newpage
\item Determine the values of $a$ for which the following system of linear equations has no solution, a unique solution, or infinitely many solutions. For each case, you can answer ``Always'', ``$a=\underline{\phantom{aaaaaaa}}$'', or ``$a\neq\underline{\phantom{aaaaaaa}}$''.
\[
 \begin{array}{ccccccc}
  ax_1&-&9x_2&-&6x_3&=&0\\
-2x_1&+&6x_2&+&6x_3&=&0\\
2x_1&-&6x_2&-&7x_3&=&0
 \end{array}
\]
No Solutions:\\
\\
Unique Solution:\\
\\
Infinitely Many Solutions:
\\
\item Find the row-echelon form of the matrix $A$, and give its rank:
\[
 A=\begin{bmatrix}2&-4&4\\-1&2&3\\1&-2&-1\\3&6&-3\end{bmatrix}
\]
\item Consider the following matrices:
\[
 A = \begin{bmatrix}
      3\\2\\-2
     \end{bmatrix}\quad B = \begin{bmatrix}9\\8\\-5\end{bmatrix} \quad C = \begin{bmatrix}0\\2\\1\end{bmatrix}
\]
For each of the following matrices, determine whether it can be written as a linear combination of the matrices $A$, $B$, and $C$. If not, answer ``No''. If it can, write your linear combination using the matrix names above.
\[
 V_1 = \begin{bmatrix}-3\\4\\1\end{bmatrix}\quad V_2 = \begin{bmatrix}0\\-2\\-1\end{bmatrix}\quad V_3 = \begin{bmatrix}15\\8\\-11\end{bmatrix}
\]
\item Compute the following matrix sum:
\[
 \begin{bmatrix}
  -6&0\\5&10\\-10&3\\2&-7\\0&-6
 \end{bmatrix} + \begin{bmatrix}2&10\\-1&-1\\6&2\\-1&0\\5&3\end{bmatrix} = 
\]
\item Compute the following scalar multiple:
\[
 -3\begin{bmatrix}-4&-1&0&-5&3\\-9&10&1&-9&-9\end{bmatrix}=
\]
\item Express the following system of linear equations as a vector equation and as a matrix equation:
\[
 \begin{array}{ccccc}
  -6x_1&-&6x_2&=&5\\
-6x_1&-&10x_2&=&-6
 \end{array}
\]
Number of variables:\\

Vector equation: $x_1\begin{bmatrix}?\\?\end{bmatrix} +x_2\begin{bmatrix}?\\?\end{bmatrix}= \begin{bmatrix}?\\?\end{bmatrix}$

Matrix equation: $\begin{bmatrix}?&?\\?&?\end{bmatrix}\begin{bmatrix}x_1\\x_2\end{bmatrix} = \begin{bmatrix}?\\?\end{bmatrix}$


 \end{enumerate}
\end{document}
 
