\documentclass[letterpaper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{enumerate}

\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\inter}[1]{\overset{\,\,\circ}{#1}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\cla}[1]{\left[ #1\right]}
\DeclareMathOperator{\Int}{Int}

\title{Math 4310 Assignment \#11\\University of Lethbridge, Fall 2014}
\author{Sean Fitzpatrick}
\begin{document}
 \maketitle

{\bf Due date:} Friday, November 21st, by 5 pm.

\bigskip

Please submit solutions to the following problems (except where indicated). As usual, all the textbook exercises are recommended as practice problems if they don't appear as an assigned problem below.

\begin{enumerate}
\item \begin{enumerate}
\item Let $\gamma: [0,1]\to X$ be a path, and let $\rho:[0,1]\to [0,1]$ be any continuous function such that $\rho(0)=0$ and $\rho(1)=1$. Prove that the paths $\gamma$ and $\gamma\circ \rho$ are homotopic.

Hint: $\rho$ is itself a path from 0 to 1 in $[0,1]$, and all such paths are homotopic in $[0,1]$.
\item Let $\alpha, \beta$, and $\gamma$ be loops based at a point $x_0\in X$. Write down explicit formulas for $\alpha\ast(\beta\ast\gamma)$ and $(\alpha\ast\beta)\ast\gamma$.
\item Prove that $\cla{\alpha}\ast(\cla{\beta}\ast\cla{\gamma}) = (\cla{\alpha}\ast\cla{\beta})\ast\cla{\gamma}$.

Hint: use (a), and try the map $\displaystyle \rho(s)=\begin{cases} s/2 & \text{ if } 0\leq s\leq 1/2\\ s-1/4 & \text{ if } 1/2\leq s\leq 3/4\\ 2s-1 & \text{ if } 3/4\leq s\leq 1\end{cases}$.

Note: A similar approach can be used to verify the other group axioms as well. For example, let $e_{x_0}:[0,1]\to X$ be the constant map $e_{x_0}(s)=x_0$ for all $s\in [0,1]$, and let $\gamma:[0,1]\to X$ be a loop based at $x_0$. Let $e_0:[0,1]\to [0,1]$ be the constant map $e_0(t)=0$ for all $t\in [0,1]$, and let $i:[0,1]\to [0,1]$ be the identity map. Then $e_0\ast i:[0,1]\to [0,1]$ is homotopic to $i$, and 
\[
\gamma\circ(e_0\ast i) = (\gamma \circ e_0)\ast (\gamma \circ i) = e_{x_0}\ast \gamma.
\]
If $F$ is a homotopy from $e_0\ast i$ to $i$, then $\gamma\circ F$ is a homotopy from $e_{x_0}\ast \gamma$ to $\gamma\circ i = \gamma$. Thus, $\cla{e_{x_0}}\ast\cla{\gamma} = \cla{\gamma}$.


(Here we used the fact that for any paths $\alpha$ and $\beta$ in a space $Y$ and a map $f:Y\to X$, $f\circ(\alpha\ast\beta) = (f\circ \alpha)\ast(f\circ \beta)$ since both are equal to the map given by $f(\alpha(2s))$ for $s\in [0,1/2]$ and $f(\beta(2s-1))$ for $s\in [1/2,1]$, and as mentioned in class, if $F$ is a homotopy from $\alpha$ to $\beta$ in $Y$, then $G=f\circ F$ is a homotopy from $f\circ \alpha$ to $f\circ \beta$ in $X$.)
\end{enumerate}
\item Let $X$ be a space and let $\alpha,\beta:[0,1]\to X$ be two paths from $x_0$ to $x_1$, for two points $x_0,x_1\in X$. These paths define isomorphisms $\varphi_\alpha,\varphi_\beta:\pi_1(X,x_0)\to \pi_1(X,x_1)$, but as noted in class, they may be different isomorphisms. Prove that the isomorphism $\varphi_\beta$ is the composition of $\varphi_\alpha$ with the inner automorphism of $\pi_1(X,x_1)$ induced by the element $\cla{\beta^{-1}\ast\alpha}$.
\item Prove that the two isomorphisms in the previous problem are the same if and only if $\pi_1(X,x_0)$ is Abelian.
\item Given spaces $X$ and $Y$, let $\cla{X,Y}$ denote the set of homotopy classes of maps $f:X\to Y$.
\begin{enumerate}
\item Let $I=[0,1]$. Show that for any space $X$, $\cla{X,I}$ contains a single element.
\item Show that if $Y$ is path connected, then the set $\cla{I,Y}$ contains a single element.
\end{enumerate}
\item ({\bf Do not submit}) A space $X$ is called {\bf contractible} if the identity map $i_X:X\to X$ is homotopic to a constant map. (If $f$ is homotopic to a constant map, we say $f$ is {\bf nullhomotopic}.)
\begin{enumerate}
\item Show that $I$ and $\R$ are contractible.
\item Show that a contractible space is path-connected.
\item Show that if $Y$ is contractible, then for any set $X$, the set $\cla{X,Y}$ has a single element.
\item Show that if $X$ is contractible and $Y$ is path-connected, then the set $\cla{X,Y}$ has a single element.
\end{enumerate}
\item Let $A\subseteq X$. Recall that a {\bf retraction} of $X$ onto $A$ is a continuous map $r:X\to A$ such that $r(a)=a$ for all $a\in A$. If $a_0\in A$, show that
\[
r_*:\pi_1(X,a_0)\to \pi_1(A,a_0)
\]
is a surjection.
\end{enumerate}
\end{document}
 
