\documentclass[letterpaper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{enumerate}

\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\inter}[1]{\overset{\,\,\circ}{#1}}
\newcommand{\T}{\mathcal{T}}
\DeclareMathOperator{\Int}{Int}

\title{Math 4310 Assignment \#8\\University of Lethbridge, Fall 2014}
\author{Sean Fitzpatrick}
\begin{document}
 \maketitle

{\bf Due date:} Friday, October 31st, by 5 pm.

\bigskip

Please submit solutions to the following problems (except where indicated). As usual, all the textbook exercises are recommended as practice problems if they don't appear as an assigned problem below.

\begin{enumerate}
\item Prove that any finite subset of a topological space is compact.
\item Let $X$ be a set and let $\T_1, \T_2$ be two topologies on $X$, such that $\T_1\subseteq \T_2$.
\begin{enumerate}
 \item Prove that if $(X,\T_2)$ is compact, then $(X,\T_1)$ is compact.
 \item Prove that if $(X,\T_1)$ is Hausdorff and $(X,\T_2)$ is compact, then $\T_1=\T_2$.
\end{enumerate}
 \item Prove that if $\{A_\alpha\}$ is any collection of compact subsets of a Hausdorff space $X$, then $\bigcap_\alpha A_\alpha$ is compact.
 \item Prove that if $Y$ is compact, then the projection $\pi_X: X\times Y\to X$ is a closed map.
 \item Prove the following theorem: Let $Y$ be a compact Hausdorff space, and let $f:X\to Y$ be a map. Then $f$ is continuous if and only if the graph of $f$, $\Gamma_f = \{(x,f(x)) : x\in X\}$ is closed in $X\times Y$.

{\em Hint:} If $\Gamma_f$ is closed and $V$ is a neighbourhood of $f(x_0)$ in $Y$, then the intersection of $\Gamma_f$ and $X\times (Y\setminus V)$ is closed. Now apply the previous problem.
\end{enumerate}
\end{document}
 
