\documentclass[letterpaper,12pt]{amsart}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{multicol}
\usepackage[bitstream-charter]{mathdesign}
\usepackage[T1]{fontenc}

\newcommand{\len}[1]{\lVert #1\rVert}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}
                   
\title{Math 1410 Assignment \#5\\University of Lethbridge, Fall 2016}
\author{Sean Fitzpatrick}
\begin{document}
 \maketitle

{\bf Due date:} {\bf Tuesday}, November 29th, by 4:00 pm.

\bigskip

Please review the {\bf Guidelines for preparing your assignments} before submitting your work. You can find these guidelines, along with the required cover page, in the Assignments section on our Moodle site.



\subsection*{Assigned problems}
\begin{enumerate}

\item Recall that an $n\times n$ matrix $A$ is {\bf idempotent} if $A^2=A$. Show that:

\medskip

\begin{enumerate}
 \item The identity matrix $I$ is the only invertible idempotent matrix.

\medskip

 \item A matrix $A$ is idempotent if and only if $I-2A$ is self-inverse. 

\medskip

 \item If $A$ is idempotent, then $I-kA$ is invertible for any $k\neq 1$, and
\[
 (I-kA)^{-1} = I+\left(\frac{k}{1-k}\right)A.
\]
\end{enumerate}

\bigskip


\item Recall that an $n\times n$ matrix $A$ is {\bf symmetric} if $A^T=A$, and {\bf antisymmetric} if $A^T=-A$.

\bigskip

\begin{enumerate}
 \item Show that $B+B^T$ is symmetric for {\bf any} $n\times n$ matrix $B$.

\medskip

 \item Show that $B-B^T$ is antisymmetric for {\bf any} $n\times n$ matrix $B$.

\medskip

 \item Given any $n\times n$ matrix $B$, find a symmetric matrix $U$ and an antisymmetric matrix $V$ such that $B=U+V$.
\end{enumerate}

\bigskip

\item What can be said about the determinant of $A$ if:

\medskip

\begin{enumerate}
 \item $A$ is idempotent.

\medskip

 \item $A$ is self-inverse.

\medskip

 \item $A$ is antisymmetric.  
\end{enumerate}

\bigskip

\item Determine all values of $k$ such that the following matrices are invertible:

\medskip

\[
 A = \bbm k&-k&3\\0&k+1&1\\k&-8&k-1\ebm \quad B = \bbm k&k&0\\k^2&4&k^2\\0&k&k\ebm
\]

\end{enumerate}

\bigskip

\bigskip

{\bf See over} for hints and suggestions.

\newpage
\textbf{Hints and suggestions:}

\bigskip

\begin{itemize}
 \item For 1(b), to prove an ``A if and only if B'' statement you must prove two things: A implies B, and B implies A. In this case that means you must prove (i) if $A$ is idempotent, then $I-2A$ is self-inverse, and (ii) if $I-2A$ is self-inverse, then $A$ is idempotent.

\medskip
 \item For 1(c), remember that showing $B=A^{-1}$ is the same thing as showing $AB=I$.

\medskip
 \item For 2(a) and 2(b) you should be giving a \textit{general} argument, relying only on the properties of the transpose. Your results here are intended as a hint for 2(c).

\medskip

 \item For problem 3(c), assume $A$ is an $n\times n$ matrix. There are two cases, depending on whether $n$ is even or odd.

\medskip

 \item For problem 4, use determinants!
\end{itemize}

\end{document}
 
