\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 2000}\chead{Quiz \# 10}\rhead{Thursday, 20\textsuperscript{th} November, 2014}
\rfoot{Total: 10 points}

\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\di}{\displaystyle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\modd}[3]{#1\equiv #2 \pmod{#3}}

\begin{document}

%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
\noindent{{\bf Name and student number:}}

 \begin{enumerate}
 \item \begin{enumerate}
        \item Suppose $k\leq l$ and let $\N_k=\{1,2,\ldots, k\}$ and let $\N_l = \{1,2,\ldots, l\}$. Show that the function $f:\N_k\to \N_l$ given by $f(n) = n$ for $n=1,2,\ldots, k$ is one-to-one.\points{2}

\vspace{2in}

	\item Is the function $f$ in part (a) necessarily onto? Explain. \points{1}

\vspace{1.5in}

	\item Suppose $k\leq l$, and suppose $A$ and $B$ are sets, with $\abs{A}=k$ and $\abs{B}=l$. Let $g:A\to \N_k$ and $h:B\to \N_l$ be bijections. (The bijections $g$ and $h$ exist, by the definition of the cardinality of finite sets.) Use part (a) and the bijections $g$ and $h$ to construct a one-to-one function from $A$ to $B$. \points{2}
       \end{enumerate}

\newpage

\item Suppose every student at a university has three initials, say F.M.L. for First, Middle, Last. How many students must the university have to guarantee that two students have the same initials? \points{5}

{\em Hint}: Let $A$ be the set of university students, and let $B$ be the set of letters in the alphabet, so $\abs{B}=26$. The set of all possible initials can be identified with $B\times B\times B$, since an ordered triple of letters $(b_1,b_2,b_3)$ corresponds to a set of initials. If $f:A\to B\times B\times B$ is the function that assigns each student to their initials, how big must the cardinality of $A$ be to guarantee that $f$ cannot be one-to-one?

 \end{enumerate}
\end{document}