\documentclass[letterpaper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{multicol}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\modd}[3]{#1\equiv #2 \pmod{#3}}

\title{Practice Problems for Quiz 6\\Math 2000A}
\date{}
\begin{document}
 \maketitle
\vspace{-0.5in}

Quiz \#6 will take place in class on Thursday, October 23rd. 

Feel free to discuss the solutions to these quiz problems on Piazza. (As a reminder, this earns you participation credit.)
\begin{enumerate}
 \item Determine if each of the assertions below is true or false. If the assertion is true, provide a proof. If not, provide a counterexample.
\begin{enumerate}
 \item For all integers $a$, $b$, and $c$, with $a\neq 0$, if $a|b$ and $a|c$, then $a|(b-c)$.
 \item For each $n\in \Z$, if $n$ is odd, then $n^3$ is odd.
 \item For all integers $a$, $b$, and $c$ with $a\neq 0$, if $a|b$, then $a|(bc)$.
 \item For every integer $n$, $4n^2+7n+6$ is an odd integer.
 \item For all integers $a$, $b$, and $d$ with $d\neq 0$, if $d$ divides both $a-b$ and $a+b$, then $d$ divides $a$.
\end{enumerate}
 \item If $x$ and $y$ are integers, explain why $xy=1$ implies either $x=1$ or $x=-1$.
 \item Is the following true or false? For all nonzero integers $a$ and $b$, if $a|b$ and $b|a$, then $a=\pm b$.
 \item Let $a$ and $b$ be integers. Prove that if $\modd{a}{2}{3}$ and $\modd{b}{2}{3}$, then $\modd{a+b}{1}{3}$ and $\modd{ab}{1}{3}$.
 \item Are the following statements true or false? Justify your conclusions.
\begin{enumerate}
 \item For each $a\in\Z$, if $\modd{a}{2}{5}$, then $\modd{a^2}{4}{5}$.
 \item For each $a\in\Z$, if $\modd{a^2}{4}{5}$, then $\modd{a}{2}{5}$.
 \item For each $a\in\Z$, $\modd{a}{2}{5}$ if and only if $\modd{a^2}{4}{5}$.
\end{enumerate}
\item Are the following statements true or false? Justify your conclusions. (For true statements, you might find that proof by contradiction is useful.)
\begin{enumerate}
 \item For all integers $a$ and $b$, if $a$ is even and $b$ is odd, then 4 does not divide $a^2+b^2$.
 \item For all integers $a$ and $b$, if $a$ is even and $b$ is odd, then 6 does not divide $a^2+b^2$.
 \item For all integers $a$ and $b$, if $a$ is even and $b$ is odd, then 4 does not divide $a^2+2b^2$.
 \item For all integers $a$ and $b$, if $a$ is odd and $b$ is odd, then 4 divides $a^2+3b^2$.
\end{enumerate}
\newpage

\item \begin{enumerate}
 \item Prove that for each $a\in\Z$, $a\not\equiv 0\pmod{3}$ if and only if $\modd{a^2}{1}{3}$
 \item Prove that for each $n\in\N$, $\sqrt{3n+2}$ is not a natural number. 
\end{enumerate}
\item Let $A$, $B$, and $C$ be sets. Establish the following:
\begin{enumerate}
 \item $A\times (B\cap C) = (A\times B)\cap (A\times C)$
 \item $A\times (B\cup C) = (A\times B)\cup (A\times C)$
 \item $(A\cap B)\times C = (A\times C)\cap (B\times C)$
 \item $(A\cup B)\times C = (A\times C)\cup (B\times C)$
 \item $A\times (B\setminus C) = (A\times B)\setminus (A\times C)$
 \item $(A\setminus B)\times C = (A\times C)\setminus (B\times C)$
 \item If $S\subseteq A$, then $S\times B\subseteq A\times B$
 \item If $T\subseteq B$, then $A\times T\subseteq A\times B$
\end{enumerate}
\item Let $A=\{1\}$, $B=\{2\}$, and $C=\{3\}$.
\begin{enumerate}
 \item Explain why $A\times B\neq B\times A$.
 \item Explain why $(A\times B)\times C \neq A\times (B\times C)$.
\end{enumerate}
 \item Let $f:(\R\setminus\{0\})\to \R$ be the function defined by $f(x)=\dfrac{x^3+5x}{x}$, and let $g:\R\to\R$ be the function defined by $g(x)=x^2+5$.
\begin{enumerate}
 \item Calculate $f(2), f(-2), f(3)$, and $f(\sqrt{2})$.
 \item Calculate $g(2), g(-2), g(3)$, and $g(\sqrt{2})$.
 \item Is the function $f$ equal to the function $g$? Explain.
 \item If we let $h:(\R\setminus \{0\})\to \R$ be the function defined by $h(x)=x^2+5$, is the function $h$ equal to $f$? Explain.
\end{enumerate}






\end{enumerate}


\end{document}
 
