\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 3410A}\chead{Quiz \# 4}\rhead{Friday, 6\textsuperscript{th} February, 2015}
\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}

\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\range}{range}
\begin{document}

%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}
{\bf Name: Solutions}

\bigskip

Solve {\bf one} of the following two questions:
 \begin{enumerate}
 \item  Suppose $T\in\mathcal{L}(V,W)$ and $v_1,v_2,\ldots, v_m$ are vectors in $V$ such that the vectors $Tv_1,Tv_2,\ldots, Tv_m$ are linearly independent in $W$. Prove that the vectors $v_1,\ldots, v_m$ are linearly independent in $V$. 

\bigskip

\noindent{\bf Solution:} Suppose that the vectors $Tv_1,\ldots, Tv_m$ are linearly independent. We want to show that the vectors $v_1,\ldots, v_m$ are linearly independent. Thus, we suppose that
\[
 c_1v_1+c_2v_2+\cdots + c_mv_m = 0
\]
for some scalars $c_1,c_2,\ldots, c_m\in\mathbb{F}$. We need to show that we're forced to take each of these scalars equal to zero. Since $T$ is a linear transformation, we know that $T(0)=0$. Therefore, we have
\begin{align*}
 0& = T(0)\\
 & = T(c_1v_1+c_2v_2+\cdots +c_mv_m) &\text{ (since  } c_1v_1+c_2v_2+\cdots + c_mv_m = 0)\\
 & = c_1Tv_1+c_2Tv_2+\cdots +c_mTv_m &\text{ (since } T \text{ is a linear map)}
\end{align*}
But this means we have a linear combination of the vectors $Tv_1,\ldots, Tv_m$ equal to zero, and these vectors were assumed to be linearly independent. Thus, the only possibility is that each of the scalars is zero: $c_1=0,c_2=0,\ldots, c_m=0$. But this is what we needed to show. Therefore, the vectors $v_1,\ldots, v_m$ are linearly independent.


\bigskip

\item Suppose that the vectors $v_1,\ldots, v_m$ span the vector space $V$, and that $T:V\to W$ is a linear transformation. Prove that the vectors $Tv_1,\ldots, Tv_m$ span $\range T$.

\bigskip

\noindent{\bf Solution:} Suppose that $V=\spn\{v_1,\ldots, v_m\}$, and that $w\in \range T$. By definition, if $w\in \range T$, then there is some $v\in V$ such that $Tv=w$. Since we know that $V=\spn\{v_1,\ldots, v_m\}$ and $v\in V$, it follows that there exist scalars $c_1,\ldots, c_m\in\mathbb{F}$ such that
\[
 v = c_1v_1+c_2v_2+\cdots + c_mv_m.
\]
Using the linearity of $T$, this implies that
\begin{align*}
 w & = Tv\\
 & = T(c_1v_1+c_2v_2+\cdots + c_mv_m)\\
 & = c_1Tv_1+c_2Tv_2+\cdots +c_mTv_m.
\end{align*}
But this means that any $w\in\range T$ can be written as a linear combination of the vectors $Tv_1,Tv_2,\ldots, Tv_m$, which is exactly the definition of what it means to say that the vectors $Tv_1,Tv_2,\ldots, Tv_m$ span the range of $T$, so we're done.

\bigskip

 \end{enumerate}
\end{document}