\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,top=1.5in,bottom=1in,left=0.75in,right=0.75in,centering]{geometry}
%\usepackage{fancyhdr}
\usepackage{enumerate}
%\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

%\pagestyle{fancy}
%\cfoot{}
%\lhead{Math 1560}\chead{Test \# 1}\rhead{May 18th, 2017}
%\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\len}[1]{\lVert #1\rVert}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}                   
                  
\begin{document}


\author{Instructor: Sean Fitzpatrick}
\thispagestyle{empty}
\vglue1cm
\begin{center}
\emph{University of Lethbridge}\\
Department of Mathematics and Computer Science\\
{\bf MATH 1560 - Tutorial \#7}\\
Monday, March 5
\end{center}
\skipline \ \noindent \skipline

\vspace*{\fill}


Some additional practice (discuss the answers but don't write anything down):
\begin{enumerate}
\item A person standing on a ledge $75$ m above the ground throws a ball upward at $20$ m/s. Assume that acceleration due to gravity is $-10\text{ m/s}^2$.
\begin{enumerate}
\item How high (from the ground) does the ball get?
\item How many seconds does it take before the ball hits the ground?
\end{enumerate}
\item A baseball diamond is a square with sides of length 90 feet. A baseball player hits the ball, and runs toward first base at a speed of 24 ft/s. 
\begin{enumerate}
\item At what rate is his distance from second base decreasing when he is halfway to first base?
\item At what rate is his distance from third base increasing at the same moment?
\end{enumerate}
\end{enumerate}




\newpage
%\thispagestyle{empty}


\begin{enumerate}
\item An aircraft is flying away from a viewer on the ground at a speed of $210 \text{ m/s}$ at an elevation of $130 \text{ m}$. The person on the ground is watching the plane through a set of binoculars. Let $\theta$ denote the viewing angle, measured from the ground, in radians. At what rate is $\theta$ changing when the plane is $500 \text{ m}$ away (as measured along the ground)?

\vspace{3.5in}

\item Suppose a police officer is 300 metres south of an intersection, and travelling north at 80 km/h. At the same time, a vehicle is 400 metres east of the same intersection, travelling east (away from the intersection). 

If the officer's radar gun registers a speed of 32 km/h, how fast is the car travelling?

\end{enumerate}
\end{document}