\documentclass[letterpaper,12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}

\usepackage{hyperref}

\author{Sean Fitzpatrick}



%opening
\title{Guidelines for preparing your assignments}
\author{Math 1410\\Spring 2017\\Sean Fitzpatrick}

\begin{document}

\maketitle

\begin{itemize}
 \item Assignments should be submitted to the appropriate Math 1410 \textbf{assignment drop box}, located across from the Math \& CS Department office in C526.
 \item You must attach a cover page. The assignment cover page template is available on Moodle. Print it off, and fill it out.
 \item Assignments may be completed individually, or in groups of up to four students. There is one version of the cover page available for each option.
 \item For group assignments, it is up to the members of the group to ensure that the workload is shared fairly. If someone isn't contributing, I don't want to know about it. Tell them to submit their own work and refuse to let them copy yours.
 \item Late assignments are not accepted without {\bf prior} permission. 
 \item Plagiarism is not tolerated. This includes copying from your friend and using online resources without citation. If you receive help from a classmate or a tutor, you should cite them as a resource on your cover page.
 \item Remarks on plagiarism:
\begin{itemize}
 \item It's perfectly OK to use a website or software to help you solve a problem, as long as you cite it as a resource. For example, if you're asked to solve a system of equations (and you're not told that you have to write out all the steps), it's perfectly fine to write something like, ``Using Wolfram Alpha, we found the solution to be....'' However, if you were to write down that solution with no citation and no work to show how you obtained the answer, we're going to get suspicious.
 \item If you \textbf{are} willing to let your friend copy, save us all some trouble and just use the group cover page, and list your friend as a group member.
 \item In cases where plagiarism is detected, the standard penalty is a zero on the assignment, and a letter to the Dean's office placing the offence on your permanent record.
\end{itemize}
\pagebreak
 \item Assignments must be professionally prepared. Expectations include:
 \begin{itemize}
  \item Inclusion of the cover page.
  \item Stapling your assignment (one staple, upper left-hand corner). There's a stapler next to the drop box, so there's no excuse for not stapling your assignment.
  \item A \textbf{good copy}. This means nothing should be scribbled out, and nothing should be spilled on the pages (food, coffee, tears, etc.). Paper torn from a notebook is not acceptable (especially if there are frayed edges).
  \item Your good copy does \textbf{not} have to be typed, but it does have to be \textit{legible}. The graders will not award points for work they cannot read.
  \item Solutions should be written using \textbf{complete English sentences}. You should explain your work sufficiently clearly that anyone in the class (including yourself, one month after completing the assignment) can read your solutions and understand what you did.
 \end{itemize}
  \item The graders have permission to deduct up to 10\% of the total grade for any failures to meet the expectations above. (Multiple deductions are possible: for example, an assignment on frayed notebook paper ($-10\%$) with no cover page ($-10\%$) would lose 20\% of the total grade.)
  \item If you require clarification on any of these guidelines, please see me prior to submitting your work.
\end{itemize}

\end{document}
