\documentclass[letterpaper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
\usepackage[margin=1in]{geometry}

\usepackage[bitstream-charter]{mathdesign}
\usepackage[T1]{fontenc}

\newcommand{\len}[1]{\lVert #1\rVert}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}
\DeclareMathOperator{\proj}{proj}

\title{Math 1410 Assignment \#4\\University of Lethbridge, Spring 2017}
\author{Sean Fitzpatrick}
\begin{document}
 \maketitle

{\bf Due date:} {\bf Thursday}, March 23rd, by 4 pm.

\bigskip

Please review the {\bf Guidelines for preparing your assignments} before submitting your work. You can find these guidelines, along with the required cover page, in the Assignments section on our Moodle site.



\subsection*{Assigned problems}
\begin{enumerate}
\item Determine the null space and column space of the matrix $A = \bbm 2&-3&1&4\\-1&2&2&-3\\1&0&8&-1\ebm$.


\item Factor the matrix $A = \bbm 1&-2&3\\2&-3&1\\-1&2&4\ebm$ as a product of elementary matrices.

\item For each statement below, either prove the statement or give a counterexample showing that it is false.
\begin{enumerate}
 \item If $A$ and $B$ are both invertible, then $A+B$ is invertible.
 \item If $AB=I$, then $AB=BA$.
 \item If $AB=B$ for some matrix $B\neq 0$, then $A$ is invertible. 
 \item If $A^3$ is invertible, then $A$ is invertible.
\end{enumerate}

\item Let $A$ be a non-zero $n\times n$ matrix, and let $I$ be the $n\times n$ identity matrix.
\begin{enumerate}
 \item Show that if $A^2=0$, then $(I-A)^{-1} = I+A$.
 \item Show that if $A^3=0$, then $(I-A)^{-1} = I+A+A^2$.
 \item Find the inverse of $B = \bbm 1&3&-2\\0&1&4\\0&0&1\ebm$.
 \item Given that $A\neq 0, A^2\neq 0, \ldots, A^{n-1}\neq 0$ but $A^n=0$, determine a formula for $(I-A)^{-1}$, and show that your answer is correct.
\end{enumerate}


\end{enumerate}

\end{document}
 
