\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{}
\lhead{Math 2000B}\chead{Quiz \# 4}\rhead{Thursday, 5\textsuperscript{th} February, 2015}
\rfoot{Total: 10 points}
%\chead{{\bf Name:}}
\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\renewcommand{\div}[2]{#1\, |\,#2}

\begin{document}
{\bf Name: Solutions}
%\author{Instructor: Sean Fitzpatrick}
\thispagestyle{fancy}
%\noindent{{\bf Name and student number:}}

\bigskip
Choose {\bf one} of the following two problems:
\begin{enumerate}
 \item Prove the following statement: For all nonzero real numbers $x$ and $y$, if $x$ is a rational number and $y$ is an irrational number, then $\dfrac{x}{y}$ is an irrational number.
     

\begin{proof}
Suppose to the contrary that $x$ is rational and $y$ is irrational, but $\dfrac{x}{y}$ is rational. Since $x\neq 0$, $\dfrac{y}{x}$ is defined, and is rational, since it's equal to $\dfrac{1}{x/y}$. But if $x$ and $\dfrac{y}{x}$ are both rational, then
\[
 y = x\left(\dfrac{y}{x}\right)
\]
 is rational, contradicting the assumption that $y$ is irrational. Thus, it must be the case that $\dfrac{x}{y}$ is rational.
\end{proof}

\item Prove the following statement: For all real numbers $x$, either $x+\sqrt{2}$ is irrational, or $-x+\sqrt{2}$ is irrational.

\begin{proof}
 Suppose to the contrary that $x+\sqrt{2}$ and $-x+\sqrt{2}$ are both rational. If this is the case, then it's also true that $a = \frac{1}{2}(x+\sqrt{2})$ and $b=\frac{1}{2}(-x+\sqrt{2})$ are rational.

Since the sum of two rational numbers is rational, this implies that
\[
 a+b = \frac{x}{2}+\frac{\sqrt{2}}{2}-\frac{x}{2}+\frac{\sqrt{2}}{2} = \sqrt{2}
\]
is rational, which contradicts the fact that $\sqrt{2}$ is irrational. Since we arrived at our contradiction by assuming that both $x+\sqrt{2}$ and $-x+\sqrt{2}$ are rational, it must be the case that at least one of them is irrational.
\end{proof}
Note: for \#2, if $P(x)$ is the predicate ``$x+\sqrt{2}$ is irrational'' and $Q(x)$ is the predicate ``$-x+\sqrt{2}$ is irrational'', then we are trying to prove $P(x)\vee Q(x)$, and we start our proof by contradiction by assuming the negation, which is $\neg P(x)\wedge \neg Q(x)$, by one of de Morgan's laws.

\end{enumerate}



\end{document}