\documentclass[letterpaper,12pt,landscape]{amsart}

%\usepackage{ucs}
%\usepackage[utf8x]{inputenc}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage[canadian]{babel}
\usepackage[margin=0.5in]{geometry}
\usepackage{tabularx}
%\usepackage[dvips]{hyperref}

\author{Sean Fitzpatrick}
\title{Grading Rubric for Assignment \#1\\Math 2000 A \& B}

\begin{document}
 \maketitle

Since a number of people have asked for more details regarding the first assignment, I've put together a grading rubric that I'll use to grade the assignments, which I've provided below. Note that it is not necessary to type your assignment, as long as your handwriting is legible. (Formatting equations can be time-consuming.) The length of the assignment will depend on the size of your writing and how consise your solutions to the problems are, but you should probably end up with something around 3-5 pages.

The grading criteria below are based on the instructions on the assignment, which you can access on Moodle.

\begin{center}
 \begin{tabularx}{\textwidth}{|c|X|X|X|X|X|c|c|}
\hline
   & \multicolumn{5}{c|}{{\bf Grading Criteria}} & \multicolumn{2}{c|}{{\bf Points}}\\
\hline
 & {\bf Excellent} & {\bf Good} & {\bf Acceptable} & {\bf Needs Improvement} & {\bf Missing or flawed} & {\bf Total} & {\bf Score}\\
\hline
{\bf Format} & The paper is presented in paragraph form, and is easy and interesting to read.& The paper is in paragraph form but presentation (or spelling, grammar, etc.) could be improved. & Content is organized, and paper contains an introduction and conclusion. & Solutions are presented in numbered format, with no connections between them. & Solutions are poorly written and/or unorganized. & 6 & \\
\hline
{\bf Solutions} & All problems are solved correctly and proofs are well-written. & Solutions contain a few mistakes or omissions. & A few proofs are incorrect or missing. & Proofs contain logical errors or are not written using complete sentences. & Most proofs are incorrect or missing.&10 & \\
\hline
{\bf References} & Paper contains additional material from outside sources, with references. &  &  & No material from outside sources. & Paper contains plagiarized content (references missing). & 4 & \\
\hline
{\bf Total} & 90-100\% & 80-89\% & 65-79\% & 50-64\% & 0-49\% & 20 &\\
\hline
 \end{tabularx}

\end{center}

\end{document}
 
