\documentclass[letterpaper,12pt]{article}

%\usepackage{ucs}
%\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage[canadian]{babel}
\usepackage[margin=1in]{geometry}
\usepackage{multicol}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}
\newcommand{\pd}[2]{\dfrac{\partial #1}{\partial #2}}
\newcommand{\di}{\displaystyle}

\title{Practice problems for change of variables, including Quizzes 14 and 15\\Math 2580\\Spring 2016}
\author{Sean Fitzpatrick}
\date{March 1st, 2016}

\begin{document}
 \maketitle


For Quiz 14 on Tuesday, you should be able to do the following:

\begin{enumerate}
 \item Let $D$ be the region bounded by the polar curve $r=2\sin\theta$.
 \begin{enumerate}
 \item Find the area of $D$. (Caution: first identify the curve, and choose your limits of integration accordingly.
 \item Find the volume of the solid that lies above the region $D$ and below the graph of the function $f(x,y) = 4-\sqrt{x^2+y^2}$.
 \end{enumerate}
 \item Evaluate $\di \iint_D e^{x^2+y^2}\,dA$, where $D$ is the region bounded by the circles $x^2+y^2=4$ and $x^2+y^2=9$.
 \item The volume of a solid $T$ is given in cylindrical coordinates by $\di \int_0^{\pi/2}\int_0^2\int_0^{4-r^2}r\,dz\,dz\,d\theta$. Sketch the solid and compute its volume.
 \item Evaluate the integral $\iiint_W \frac{1}{\sqrt{x^2+y^2}}\,dV$, where $W$ is the region given by $0\leq x\leq \sqrt{9-y^2}$, $0\leq y\leq 3$, $0\leq z\leq \sqrt{9-(x^2-y^2)}$, using cylindrical coordinates.
 \item Find the volume that the cylinder $r=a\cos\theta$ cuts out of the sphere $x^2+y^2+z^2=a^2$, where $a>0$.
 \item Calculate the Jacobian of the transformation $T(u,v) = (u^2v, uv^2)$.

\end{enumerate}

\pagebreak


For Quiz 15 on Thursday, you should be able to do the following:

\begin{enumerate}

 \item The integral $\di \int_{-2}^2\int_0^{\sqrt{4-x^2}}\int_{x^2+y^2}^{\sqrt{8-x^2-y^2}}\,dz\,dy\,dx$ represents the volume of a solid. Sketch the solid, and compute its volume using either cylindrical or spherical coordinates, whichever you prefer.
 \item  Given the region $R$ in the $xy$-plane bounded by the hyperbolas $y=1/x$, $y=4/x$, and the lines $y=x$, $y=4x$, find equations for a transformation $T$ that maps a rectangular region $S$ in the $uv$-plane onto $R$, where the sides of $S$ are parallel to the $u$- and $v$-axes.

\item  Evaluate the integral $\di \iint_R (x+y)e^{x^2-y^2}\,dA$, where $R$ is the rectangle defined by the inequalities $0\leq x-y\leq 2$, $0\leq x+y\leq 3$ by making an appropriate change of variables.

\item Evaluate the integral $\di \iint_D(6x-4y)\,dA$, where $D$ is the region bounded by the parallelogram with vertices $(0,0)$, $(2,3)$, $(4,1)$, and $(6,4)$.

\item Use the transformation $x=u^2$, $y=v^2$, $z=w^2$ to express the volume of the solid bounded by the surface $\sqrt{x}+\sqrt{y}+\sqrt{z}=1$ and the coordinate planes as a triple integral in terms of $u$, $v$, and $w$. (You do not have to evaluate the integral. (Unless you want to.))

\end{enumerate}



Extra fun:  Let $f$ be continuous on $[0,1]$ and let $R$ be the triangular region with vertices $(0,0)$, $(1,0)$, and $(0,1)$. Show that
\[
\iint_R f(x+y)\,dA = \int_0^1uf(u)\,du.
\]

\pagebreak


Additional practice problems:

\begin{enumerate}
\item Let $D$ be the region in the first quadrant bounded by the curves $y=x$, $y=2x$, $xy=1$, and $xy=2$.
\begin{enumerate}
\item Sketch the region, carefully noting all points of intersection of these curves.
\item Determine a change of variables that transforms a rectangle in the $(u,v)$-plane into the region $D$.
\item Use this change of variables to evaluate the integral $\di \iint_D\left(\frac{(x-y)^2}{x^2}-1\right)\,dA$.
\end{enumerate}
\item Repeat the problem above, if $D$ is bounded by the curves $y=x^2$, $y=2x^2$, $x=y^2$, and $x=4y^2$, for the integral $\di \iint_D\left(\frac{x^2}{y^4}+\frac{y^2}{x^4}\right)\,dA$.
\item Using an appropriate change of variables, show that the area enclosed by the ellipse $\dfrac{x^2}{a^2}+\dfrac{y^2}{b^2}=1$ is $\pi ab$.
\item Evaluate $\iint_D \sqrt{x^2+y^2}\,dA$, if $D$ is bounded by the $y$-axis and the curve $y=\sqrt{4-x^2}$.
\item Evaluate the following integrals, either by converting to polar coordinates, or by reversing the order of integration:
\begin{enumerate}
\item $\di \int_0^4\int_{\sqrt{x}}^2 \frac{1}{y^3+1}\,dy\,dx$.
\item $\di \int_0^1\int_y^{\sqrt{2-y^2}}(x+y)\,dx\,dy$.
\item $\di \int_{-1}^1\int_0^{\sqrt{1-x^2}}\sqrt{1-y^2}\,dy\,dx$.
\end{enumerate}
\end{enumerate}
\end{document}
 
