\documentclass[letterpaper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
\newcommand{\aaa}{\mathbf{a}}
\newcommand{\bbb}{\mathbf{b}}
\newcommand{\ccc}{\mathbf{c}}
\renewcommand{\r}{\mathbf{r}}
\newcommand{\n}{\mathbf{n}}
\renewcommand{\i}{\mathbf{i}}
\renewcommand{\j}{\mathbf{j}}
\renewcommand{\k}{\mathbf{k}}

\title{Math 2580 Assignment \#7\\University of Lethbridge, Spring 2016}
\author{Sean Fitzpatrick}
\begin{document}
 \maketitle

{\bf Due date:} Thursday, April 7th, by 3 pm.

\bigskip

Please provide solutions to the problems below, using the following guidelines:
\begin{itemize}
\item Your submitted assignment should be a {\bf good copy} -- figure out the problems first, and then write down organized solutions to each problem. 
\item You should include a {\bf cover page} with the following information: the course number and title, the assignment number, your name, and a list of any resources you used or people you worked with.
\item Since you have plenty of time to work on the problems, assignment solutions will be held to a higher standard than on a test. Your explanations should be clear enough that any of your classmates can understand your solutions.
\item Group work is permitted, but copying is not. If you're not sure what the difference is, feel free to ask. If you get help solving a problem, you should (a) make sure you completely understand the solution, and (b) re-write the solution for your good copy by yourself, in your own words.
\item Assignments can be submitted in class, or in the designated drop box on the 5th floor of University Hall, across from the Math Department office.
\item Late assignments will not be accepted without prior permission.

\end{itemize}
\newpage


\subsection*{Assigned problems}
\begin{enumerate}
\item Verify Green's Theorem holds for the integral $\di \int_C (2x^3-y^3)\,dx+(x^3+y^3)\,dy$, where $C$ is the unit circle.
\item A vector field $\F(x,y) = P(x,y)\i+Q(x,y)\j$ in $\R^2$ can be viewed as a special case of a vector field in $\R^3$ that does not depend on $z$, with $z$-component equal to zero. With this identification,
\begin{enumerate}
 \item Show that $(\nabla\times\F)\dotp \k = \dfrac{\partial Q}{\partial x}-\dfrac{\partial P}{\partial y}$.
 \item Use part (a) to show that Green's Theorem can be written in the vector form
\[
 \int_{C} \F\dotp d\r = \iint_D (\nabla\times \F)\dotp \k\,dA,
\]
where $D\subseteq \R^2$ is a region to which Green's Theorem applies, and $C=\partial D$ is the positively-oriented boundary of $D$.
 \item Show that Green's Theorem implies the \textit{Divergence Theorem in the Plane}: 

Let $D\subseteq \R^2$ be a region to which Green's Theorem applies, and let $C=\partial D$ be its positively-oriented boundary. Let $\n$ denote the outward-pointing unit normal vector to $C$: if $\r:[a,b]\to \R^2$, $\r(t)=(x(t),y(t))$ defines a positively-oriented parameterization of $C$, then $\n$ is given by
\[
 \n = \frac{y'(t)\i-x'(t)\j}{\sqrt{(x'(t))^2+(y'(t))^2}}. \quad \text{(Verify this.)}
\]
If $\F=P\i+Q\j$ is a $C^1$ vector field on $D$, then
\[
 \int_C \F\dotp \n\,ds = \iint_D (\nabla \dotp \F)\,dA.
\]
\end{enumerate}
\item The \textbf{Laplacian}  is a differential operator $\Delta = \nabla^2$ that acts on functions $f:D\subseteq \R^n\to \R$, defined by
\[
 \Delta f = \nabla^2f = \nabla\dotp(\nabla f) = \frac{\partial^2 f}{\partial x_1^2}+\frac{\partial^2 f}{\partial x_2^2}+\cdots + \frac{\partial^2 f}{\partial x_n^2}.
\]
A $C^2$ function $f$ is called \textbf{harmonic} if $\Delta f = 0$. Harmonic functions are important in many areas of Engineering and Physics, such as heat transfer, electrodynamics, fluid flow, robotics\footnote{According to the internet.}, etc.
\begin{enumerate}
 \item Determine whether or not the functions $f(x,y)=e^x\sin y$, $g(x,y) = x^3+y^3$, $h(x,y) = x^3-3xy^2$ are harmonic.
 \item Prove that for any harmonic function $f$ defined on a region $D$ for which Green's Theorem holds, we have
\[
 \int_{\partial D}\frac{\partial f}{\partial y}\,dx - \frac{\partial f}{\partial x}\,dy = 0.
\]
 \end{enumerate}

\end{enumerate}



\end{document}
 
