\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[letterpaper,margin=0.85in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{enumerate}
\usepackage{lastpage}
\usepackage{multicol}
\usepackage{graphicx}

\reversemarginpar

\pagestyle{fancy}
\cfoot{Page \thepage \ of \pageref{LastPage}}\rfoot{{\bf Total Points: 100}}
\chead{MATH 1410}\lhead{FINAL EXAM}\rhead{24th April, 2015}

\newcommand{\points}[1]{\marginpar{\hspace{24pt}[#1]}}
\newcommand{\skipline}{\vspace{12pt}}
%\renewcommand{\headrulewidth}{0in}
\headheight 30pt

\newcommand{\di}{\displaystyle}
\newcommand{\abs}[1]{\lvert #1\rvert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\P}{\mathcal{P}}
\DeclareMathOperator{\nul}{null}
\DeclareMathOperator{\range}{range}
\DeclareMathOperator{\spn}{span}
\newcommand{\len}[1]{\lVert #1\rVert}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\dotp}{\boldsymbol{\cdot}}

\begin{document}

\author{Instructor: Sean Fitzpatrick}
\thispagestyle{plain}
\begin{center}
\emph{University of Lethbridge}\\
Department of Mathematics and Computer Science\\
24th April, 2015, 9:00 am - 12:00 pm\\
{\bf MATH 1410 - FINAL EXAM}\\
\end{center}
\skipline \skipline \skipline \noindent \skipline
Last Name:\underline{\hspace{350pt}}\\
\skipline
First Name:\underline{\hspace{348pt}}\\
\skipline
Student Number:\underline{\hspace{322pt}}\\
\skipline

\vspace{0.5in}


\begin{quote}

 
 Record your answers below each question in the space provided.    Left-hand pages may be used as scrap paper for rough work.  If you want any work on the left-hand pages to be graded, please indicate so on the right-hand page.
 
 \bigskip
 
Partial credit will be awarded for partially correct work, so be sure to show your work, and include all necessary justifications needed to support your arguments. 

\bigskip

\end{quote}


\vspace{0.25in}

For grader's use only:

\begin{table}[hbt]
\begin{center}
\begin{tabular}{|l|r|} \hline
Page&Grade\\
\hline \hline
\cline{1-2} 2 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} 3 & \enspace\enspace\enspace\enspace\enspace\enspace/8\\
\cline{1-2} 4 & \enspace\enspace\enspace\enspace\enspace\enspace/8\\
\cline{1-2} 5 & \enspace\enspace\enspace\enspace\enspace\enspace/8\\
\cline{1-2} 6 & \enspace\enspace\enspace\enspace\enspace\enspace/8\\
\cline{1-2} 7 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} 8 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} 9 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} 10 & \enspace\enspace\enspace\enspace\enspace\enspace/8\\
\cline{1-2} 11 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} 12 & \enspace\enspace\enspace\enspace\enspace\enspace/10\\
\cline{1-2} Total & \enspace\enspace\enspace\enspace\enspace\enspace/100\\
\hline
\end{tabular}

\skipline

\skipline

\skipline

\end{center}
\end{table}
\newpage


\begin{enumerate}
\item DEFINITIONS (2 points each):

\begin{enumerate}
 \item What is the {\bf rank} of a matrix? 

\vspace{1.25in}

 \item What does it mean to say that a system of linear equations is {\bf consistent}?

\vspace{1.25in}

 \item What does it mean to say that two vectors are {\bf orthogonal}?

\vspace{1.25in}

 \item What does it mean to say that a set of vectors $\{\vec{v}_1,\ldots, \vec{v}_k\}$ is {\bf linearly independent}?

\vspace{1.25in}

 \item What is an {\bf eigenvalue} of an $n\times n$ matrix $A$?


 
\end{enumerate}
\newpage

\item SHORT ANSWER -- Calculations (2 points each): You do not have to show your work.

\begin{enumerate}
 \item If $A=\begin{bmatrix}2&-1\\3&0\end{bmatrix}$ and $B=\begin{bmatrix}-3&4\\0&5\end{bmatrix}$, what is the matrix $3A-2B$?

\vspace{1.5in}

 \item Given $z=2-3i$ and $w=1+2i$, compute $z+w$ and $zw$.

\vspace{1.75in}

 \item If $\vec{v}=\begin{bmatrix}1&-2&3\end{bmatrix}^T$ and $\vec{w}=\begin{bmatrix}-2&4&1\end{bmatrix}^T$, compute $\vec{v}\times \vec{w}$.

\vspace{1.75in}

 \item Determine whether $x=2$, $y=3$, $z=-1$ is a solution to the system
\begin{align*}
 2x-3y+z&=-6\\
 -3x+y+2z&=5
\end{align*}

\end{enumerate}
\newpage

\item SHORT ANSWER -- More calculations (2 points each): You do not have to show your work.

\begin{enumerate}
 \item Given $\vec{a} = \begin{bmatrix}2&-3&4\end{bmatrix}^T$ and $\vec{b}=\begin{bmatrix}-1&4&5\end{bmatrix}^T$, determine the vector $\vec{x}$ such that $3\vec{a}+2\vec{x} = 4\vec{b}$.

\vspace{1.75in}

 \item Let $A$ and $B$ be $3\times 3$ matrices such that $\det A = -3$ and $\det B = 4$. What is the determinant of the matrix $C=2A^3B^TA^{-1}$?

\vspace{1.75in}

 \item What is the projection of the vector $\vec{v}=\begin{bmatrix}-2&5&3\end{bmatrix}^T$ onto the vector $\vec{w} = \begin{bmatrix}4&-3&0\end{bmatrix}^T$?

\vspace{1.75in}

 \item Is the vector $\vec{v} = \begin{bmatrix}2\\-3\end{bmatrix}$ an eigenvector of the matrix $A=\begin{bmatrix}7&2\\3&6\end{bmatrix}$?
\end{enumerate}
\newpage

\item SHORT ANSWER -- A little more theory this time (2 points each):
\begin{enumerate}
 \item Suppose $AX=B$ is the matrix form of a consistent system of 6 equations in 7 variables, such that the coefficient matrix $A$ has rank 4. How many parameters will there be in the general solution to the system?

\vspace{1.75in}

 \item Suppose the matrix $A$ is not invertible. Can there possibly exist a matrix $B$ such that $AB$ is invertible? Why or why not?

\vspace{1.75in}

 \item Suppose $\vec{a}$ and $\vec{b}$ are vectors in $\R^n$ for which it is known that $\vec{a}\dotp \vec{b} = 4$ and $(\vec{a}+\vec{b})\dotp \vec{b} = 13$. What is $\len{\vec{b}}$?

\vspace{1.75in} 

 \item Suppose $\lambda=0$ is an eigenvalue of $A$. Can $A$ be invertible? Why or why not?
\end{enumerate}
\newpage

\item Compute the determinant of the matrix \points{8}
\[
 A = \begin{bmatrix}4&-1&3&-1\\3&1&0&2\\0&1&2&2\\1&2&-1&1\end{bmatrix}
\]

\newpage

\item Let $A=\begin{bmatrix}1&0&-3\\0&1&4\\-2&2&15\end{bmatrix}$
\begin{enumerate}
 \item Compute the inverse of $A$. \points{5}

\vspace{4in}

 \item Write $A^{-1}$ as a product of elementary matrices. \points{3}

\vspace{1.75in}

 \item Write $A$ as a product of elementary matrices. \points{2}
\end{enumerate}
\newpage

\item \begin{enumerate}
       \item Find the equation of the plane containing the points $P=(0,1,2)$, $Q=(-1,0,4)$, and $R=(2,1,3)$.\points{5}

\vspace{3in}

 \item Find the area of the triangle whose corners are given by the three points $P,Q,R$ in part (a). \points{2}

\vspace{2in}

 \item Find the shortest distance from the point $P_0 = (4,-2,7)$ to the plane in part (a). \points{3}
      \end{enumerate}
\newpage

\item Find point of intersection of the line $\begin{bmatrix}x&y&z\end{bmatrix}^T = \begin{bmatrix}1&0&-2\end{bmatrix}^T+t\begin{bmatrix}-2&3&1\end{bmatrix}^T$ and the plane given by the equation $x-3y+2z=4$. \points{5}

\vspace{4in}

\item Find the line of intersection of the planes $x+y+z=7$ and $2x-y+4z=3$. \points{5}

\newpage

\item Given $z=\sqrt{3}-i$, find the value of $z^7$. Express your answer in the form $x+iy$. \points{4}

\vspace{4in}

\item Find all complex solutions to the equation $z^3=-1$. \points{4} Express your answers in the form $x+iy$.

\newpage

\item Let $A=\begin{bmatrix}0&1&1\\1&0&1\\1&1&0\end{bmatrix}$.\points{10} If possible, find an invertible matrix $P$ and a diagonal matrix $D$ such that $P^{-1}AP=D$.

(You do not have to compute $P^{-1}$ or compute the product $P^{-1}AP$ to show it equals $D$.)

\newpage

\item Let $A$ be an $n\times n$ matrix.
\begin{enumerate}
 \item Show that if $B$ is similar to $A$, then $\det A = \det B$. \points{3}

\vspace{2in}

 \item Suppose $\vec{u}_1,\vec{u}_2,\ldots, \vec{u}_k$ are eigenvectors of $A$ that all correspond to the same eigenvalue $\lambda$. Show that $\vec{v}\in\operatorname{span}\{\vec{u}_1,\ldots,\vec{u}_k\}$, then $\vec{v}$ is also an eigenvector corresponding to $\lambda$. \points{4}

\vspace{3in}

 \item Let $A$ be an $n\times n$ matrix of real numbers. Show that if $\lambda\in\mathbb{C}$ is an eigenvalue of $A$, then so is $\overline{\lambda}$. \points{3}

\end{enumerate}

\end{enumerate}
\newpage

{\bf Extra page for rough work. Do not remove if you want anything on this page to be graded.}




\end{document}